Current state and prospects of R-packages for the design of experiments

### 2022-06-15 paper/arxiv/arxiv.tex
arxiv tex updated + old supp.html deleted

All CRAN packages have a title, description, package connections
(suggests, depends and imports) and other meta-information in the
DESCRIPTION file. We use the text data from the title and description
(as accessed in 2022-05-23) in Section \ref{topics}.

@@

There have been at least 50 DoE packages since 2013 but most of the
downloads are concentrated in just a handful of packages. For example,
Figure \ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the
total package downloads in 2021 of 113 DoE packages (first released
prior to 2021); we can see from Figure \ref{fig:plot-lorenz} that bottom
90\% of DoE packages (in terms of total download count in 2021) only
share about 30\% of total downloads across all DoE packages -- in
another words, 70\% of the total downloads are due to 11 packages (10\%
of the DoE packages).

@@

\caption{Lorenz curve of the total download count for DoE packages in 2021. The red line corresponds to the line of perfect equality. The green region shows the area under the Lorenz curve and the red region shows the area of gap in equality.}\label{fig:plot-lorenz}
\end{figure}

@@

\caption{The distribution of the number of downloads for each DoE package by year. Packages were removed in any year if it was released in that year or later so that each download count is for the full year. The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year. In the last 6 years, the Gini index is consistent above 60\% each year indicating that most downloads are due to a relatively small number of packages. }\label{fig:download-share}
\end{figure}

The increase in the number of packages that are not highly downloaded
may mean that \textbf{\emph{there are more packages to construct niche
experimental designs}}. Some examples of these packages include
\texttt{qtlDesign}, \texttt{PwrGSD} and \texttt{Crossover} made for QTL
experiments, group sequential designs and crossover trials,
respectively. These packages would naturally have a smaller number of
potential users. Counterfactual to this, the increase could be due to
other external factors, such as an increase in the number of skilled
developers (and thus more pacakge contributions), a change in CRAN
policy or management to add packages (either to CRAN and/or task view),
and/or that new packages are still yet to amass users. While there is an
argument that low download counts are due to the low utility and/or
quality of the packages, packages in CRAN task views are selected by
expert maintainers, thus we can reasonably assume that any package
listed in CRAN task view are of a decent utility and quality.

If the downloads are reflective of the experimental designs used in
practice, \textbf{\emph{small set of packages appear to be sufficient
for most to construct the full set of designs of experiments needed in
practice}}. Packages of course evolve and the top downloaded packages
have had regular updates that may have broadened its scope from its
previous releases.

While in absolute terms the Gini index is high for
\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
task view (42.6\% to 74.1\%), the inequality is not as severe as other
CRAN task views as shown in Figure \ref{fig:fig-gini-all-ctvs}. We can
see in Figure \ref{fig:fig-gini-all-ctvs} that the Gini index is
generally increasing for DoE packages (as is generally the case for
other CRAN task views as shown in Figure S1 in the Supplementary
Materials) but most other CRAN task views have a Gini index of over
75\%. This suggests that other CRAN task views may have dominant
standards and in comparison to other topics, there are more
\textbf{\emph{diverse approaches to designing experiments}} and thus no
single DoE package is dominant. This observation however doesn't take
into account other approaches to generating experimental designs, like
the proprietary software, CycDesignN (Whittaker, Williams, and John
2022) and SAS (SAS Institute 1985), which may be widely used.

@@

\caption{The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages. The grey line shows the distribution of the Gini index across years for all other CRAN task views. See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view.}\label{fig:fig-gini-all-ctvs}
\end{figure}

@@

We can see in Figure \ref{fig:rank-over-time} that most of the top 10
ranking packages have been in the top 10 for the last 9 years with
\texttt{lhs} steadily climbing up the ranks in the last few years. It
should be noted that the download of one package can prompt a download
of another package; the most notable dependency network is
\texttt{AlgDesign} and \texttt{agricolae}, where the former is an import
for the latter. The full dependency network within the DoE packages is
shown later in Figure @ref(fig:plot-doe-network).

@@

The consistency in the top 10 ranking packages (Figure
\ref{fig:rank-over-time}) and the fact that most downloaded DoE packages
are first released more than 10 years ago (Figure
\ref{fig:release-date-vs-download}) indicates that either the existing
packages filled the needs of mass in practice or no new packages were
compelling for many to switch their practice. We do however also see
that the the top downloaded packages generally have more updates (see
Figure \ref{fig:release-date-vs-download}) so it is possible that the
packages have improved or broadened the scope of its usage.

@@

Figure \ref{fig:wordcloud-over-time} shows some common purposes of DoE
packages, based on bigrams in the package title and description. We show
only the bigrams as unigram was not insightful and there were not many
trigrams common across packages. In counting the bigrams, we processed
the text data as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We standardised the words to lower case and removed pluralization.
\item
  Multiple mentions of the same bigram within a package was counted as
  one (e.g.~\texttt{AlgDesign} mentions ``experimental design'' four
  times in the title and the description but this is counted as one).
\item
  Bigrams that consist the stop words (provided in
  \texttt{tidytext::stop\_words} in addition to other words we deemed
  irrelevant, e.g.~``provide'', ``e.g.'', ``calculate'' and so on; full
  list is shown in the code provided in the link under Section
  \ref{pkgs}) are removed.
\end{enumerate}

@@

In this section, we discuss these interface designs broadly, with
examples illustrated from the top downloaded packages (shown in Figure
\ref{fig:rank-over-time}). We exclude \texttt{ez} and
\texttt{DiceKriging} from the discussion as the former is predominately
visualisation of experimental data and the latter is about the analysis
of computer experiments in addition to belonging to the same suite of
packages as \texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015).

@@

Factorial experiments offer a challenge in the construction and
allocation of the treatment factors. This is reflected in a number of
packages that specifically addresses this challenge. These include
\texttt{DoE.base} (Grömping 2018) for full factorial and (regular and
irregular) orthogonal array designs via \texttt{fac.design()} and
\texttt{oa.design()}, respectively, \texttt{FrF2} (Grömping 2014) for
fractional 2-level factorial designs using \texttt{FrF2()},
\texttt{FrF2Large()} or \texttt{pb()} Plankett-Burnman designs (Plackett
and Burman 1946), \texttt{conf.design} (Venables 2013) for symmetric
confounded factorial designs via \texttt{conf.design()}, and
\texttt{BHH2} (Barrios 2016) to generate full or fractional 2-level
factorial design matrix via \texttt{ffDesMatrix()}. All these designs
generally requires the user to input the number of factors and if the
design is allowed to vary in the number of levels, the input of the
number of levels for each factor. The construction of some of these
designs are catalogue-based, e.g.~\texttt{oa.design()} in
\texttt{DoE.base}.

Response surface design, as illustrated by \texttt{rsm} (Lenth 2009), is
a factorial experiment but the focus of it is to analyse using response
surface modelling. The two well known response surface designs are
Box-Behnken design (Box and Behnken 1960) and central-composite designs
(Box and Wilson 1951) via \texttt{bbd()} and \texttt{ccd()} for
central-composite designs providing formula.

Saturated designs

\hypertarget{the-case-of-menu-designs}{%
\subsection{The case of menu designs}\label{the-case-of-menu-designs}}

Computer experiments, as discussed in Section \ref{topics}, which
generally involve space filling designs, have been on the rise. The
exemplar for this type of designs are \texttt{lhs} (Carnell 2022) and
\texttt{DiceDesign}. For the \texttt{lhs} package, functions like
\texttt{randomLHS()}, \texttt{optimumLHS()}, and \texttt{maximinLHS()},
require users to specify the sample size (\(n\)) and the number of
variables (\(p\)) then it generates a Latin hypercube sample, based on
different optimisation schemes (in this case, random, S optimal and
maxmin criteria, respectively; see package documentation for more
details). Similarly for the \texttt{DiceDesign}, there is a
comprehensive list of space-filling designs such as
\texttt{dmaxDesign()}, \texttt{lhsDesign()}, \texttt{straussDesign()}
and \texttt{wspDesign()} with input of \(n\) and \(p\) as before (among
extra parameters for some) that implement algorithms that either
maximise the entropy, produce a random Latin hypercube sample, based on
Strauss process and WSP algorithm, respectively. These designs all
output a \(n \times p\) matrix with values between 0 and 1.

The \texttt{AlgDesign} (Wheeler 2022) package offers three primary
functions for generating optimal designs: \texttt{optBlock()},
\texttt{optFederov()} and \texttt{optMonteCarlo()}. These functions in
general require a formula in terms of the supplied terms with the choice
of criterion (D, A and I for the latter two). The difference lies in the
search algorithm for the optimal design.

The \texttt{agricolae} package (de Mendiburu 2021) was the most
downloaded DoE packages in 2013 to 2018 (Figure
\ref{fig:rank-over-time}) and possibly prior to 2013 (data unavailable).
This package is the prime example of constructing designs based on menu
functions, e.g.~\texttt{design.crd()}, \texttt{design.rcbd()} and
\texttt{design.split()} construct a completely randomised design,
randomised complete block design and split-plot design, respectively.
Users typically supply the treatment labels (or number of treatments in
the case of \texttt{design.split()}) and the number of replications as
argument to these functions.

\hypertarget{the-case-of-sequential-designs}{%
\subsection{The case of sequential
designs}\label{the-case-of-sequential-designs}}

Another specialised design is sequential design (also called adaptive
sampling), which is best represented by \texttt{tgp} (Gramacy and Taddy
2010) -- these require prior information, which are used to inform the
next experimental design using \texttt{tgp.design()} and
\texttt{dopt.gp()} which requires users to supply candidate samples to
sub-sample from and a model or design from a prior design. Follow-up
experiments, which can also be classified as a sequential design, is
implemented by \texttt{BsMD} (Barrios 2020). In \texttt{BsMD}, the
follow-up design is determined by a model discriminant approach using
\texttt{MD()}.

@@

Through the exploratory data analysis of three data sources (package
download logs, package metadata and CRAN task views) outlined in Section
\ref{data}, we have observed that the the total download of DoE packages
are concentrated only on a handful of R-packages although these
represent a diverse set in comparison with other CRAN task views.
Furthermore, the data suggests that the field of experimental design is
the least collaborative field; out of all CRAN task views, it has one of
the least average number of authors and the lowest intra-connectivity
(any dependency on other packages within its own task view). There are a
number of limitations and short coming in our exploratory data analysis.
First, CRAN task views are volunteer maintained so some experimental
design packages may not be included in the DoE packages. Second, we only
use the RStudio CRAN mirror download, which may bias our observations.
Third, our analysis are limited to R-packages alone, although the sheer
quantity and variation of experimental design is unmatched with any
other programming languages. Finally, all our statements should be
treated as speculative rather than conclusive; the data are all
observational so no conclusive, generalisable statement are possible.
Regardless, the data driven nature of our analysis give an objective
insight into the field of experimental design.

The interface design (discussed in Section \ref{design}) reveals that
the most widely used DoE packages generally have functions that is 1) a
menu format (i.e.~the name of the function correspond to a particular
recipe of the design or optimal search algorithm) and 2) context is
often a second-thought (e.g.~only the number of factors is needed and
the package will assign pseudo factor names or an argument exists for
optional input of character vector that corresponds to the factor
names).

@@

@@

@@

@@

@@

>>>>

All CRAN packages have a title, description, package connections
(suggests, depends and imports of other packages) and other
meta-information in the DESCRIPTION file. We use the text data from the
title and description (as accessed in 2022-05-23) in Section
\ref{topics}.

@@

There are at least 50 DoE packages since 2013 but most of the downloads
are concentrated in just a handful of packages. For example, Figure
\ref{fig:plot-lorenz} shows a Lorenz curve (Lorenz 1905) for the total
package downloads in 2021 of 113 DoE packages (first released prior to
2021); we can see from Figure \ref{fig:plot-lorenz} that bottom 90\% of
DoE packages (in terms of total download count in 2021) only share about
27\% of total downloads across all DoE packages -- in another words,
73\% of the total downloads are due to 11 packages (10\% of the DoE
packages).

@@

\caption{Lorenz curve of the total download count for DoE packages in 2021. The red line corresponds to the line of perfect equality. The yellow region shows the area under the Lorenz curve and the red region shows the area of gap in equality.}\label{fig:plot-lorenz}
\end{figure}

@@

\caption{The distribution of the number of downloads for DoE packages by year. Packages were removed in any year if it was released in that year or later so that each download count is for the full year. The label on the bottom of the plot shows the Gini index for downloads and the number of packages with full year of download count in the corresponding year. In the last 6 years, the Gini index is consistent above 60\% indicating that most downloads are due to a relatively small number of packages. }\label{fig:download-share}
\end{figure}

The increase in the number of packages that are not highly downloaded
may mean that \textbf{\emph{there are more packages to construct niche
experimental designs}}. Some examples of these packages include
\texttt{qtlDesign}, \texttt{PwrGSD} and \texttt{Crossover} made for QTL
experiments, group sequential designs and crossover trials,
respectively. These packages would naturally have a smaller number of
potential users. Counterfactual to this, the increase could be due to
other external factors, such as an increase in the number of skilled
developers (and thus more package contributions), a change in CRAN
policy or management to add packages (either to CRAN and/or task view),
and/or that new packages are still yet to amass users. While there is an
argument that low download counts are due to the low utility and/or
quality of the packages, packages in CRAN task views are selected by
expert maintainers, thus we can reasonably assume that any package
listed in CRAN task view are of a decent utility and quality.

If the downloads are reflective of the experimental designs used in
practice, \textbf{\emph{small set of packages appear to be sufficient
for most to construct the full set of designs of experiments they need
in practice}}. Packages of course evolve and the top downloaded packages
have had regular updates that may have broadened its scope from its
previous releases.

While in absolute terms the Gini index is high for
\href{http://CRAN.R-project.org/view=ExperimentalDesign}{\emph{ExperimentalDesign}}
task view (42.6\% to 74.1\%), the inequality is not as severe as other
CRAN task views as shown in Figure \ref{fig:fig-gini-all-ctvs}. We can
see in Figure \ref{fig:fig-gini-all-ctvs} that the Gini index is
generally increasing for DoE packages (as is generally the case for
other CRAN task views as shown in Figure S1 in the Supplementary
Materials) but most other CRAN task views have a Gini index of over
75\%. This suggests that other CRAN task views may have dominant
standards and in comparison to other topics, there are more
\textbf{\emph{diverse approaches to designing experiments}} and thus no
single DoE package is dominant. This observation however doesn't take
into account other approaches to generate experimental designs, like say
the proprietary software, CycDesignN (Whittaker, Williams, and John
2022), which may be widely used.

@@

\caption{The points show the Gini index of the download counts by year for ExperimentalDesign task view with the color showing the number of packages. The grey lines show the line plots of the Gini index across years for all other CRAN task views. See Figure S1 in the Supplementary Material for the line graph for the Gini index across years by each CRAN task view.}\label{fig:fig-gini-all-ctvs}
\end{figure}

@@

We can see in Figure \ref{fig:rank-over-time} that most of the top 10
ranking packages have been in the top 10 for the last 9 years with
\texttt{lhs} steadily climbing up the ranks in the last few years. It
should be noted that the download of one package can prompt a download
of another package; the most notable package connection is
\texttt{AlgDesign} and \texttt{agricolae}, where the former is an import
for the latter. The full network of packages connections within the DoE
packages is shown later in Figure \ref{fig:plot-doe-network}.

@@

The consistency in the top 10 ranking packages (Figure
\ref{fig:rank-over-time}) and the fact that most downloaded DoE packages
are first released more than 10 years ago (Figure
\ref{fig:release-date-vs-download}) indicates that either the existing
packages fulfilled the needs of mass in practice or no new packages were
compelling for many to switch their practice. We do however also see
that the the top downloaded packages generally have more updates (see
Figure \ref{fig:release-date-vs-download}) so it is possible that the
packages have improved or broadened the scope of its usage.

@@

Figure \ref{fig:wordcloud-over-time} shows some common purposes of DoE
packages, based on bigrams in the package title and description. We only
show the bigrams as unigrams were not insightful and there were not many
trigrams common across packages. In counting the bigrams, we processed
the text data as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  We standardised the words to lower case and removed pluralization.
\item
  Multiple mentions of the same bigram within a package was counted as
  one (e.g.~\texttt{AlgDesign} mentions ``experimental design'' four
  times in the title and the description but this is counted as one).
\item
  Bigrams that consist the stop words are removed. The stop words are
  sourced from the lexicons in \texttt{tidytext::stop\_words} in
  addition to other words we deemed irrelevant, e.g.~``provide'',
  ``e.g.'', ``calculate'' and so on -- full list is shown in the code
  provided in the link under Section \ref{pkgs}.
\end{enumerate}

@@

In this section, we discuss the interface designs of functions that
output an experimental design based on three broad areas: factorial
designs, recipe designs and augmenting designs. The discussion is
exclusive to the top downloaded packages (shown in Figure
\ref{fig:rank-over-time}), with the exception of \texttt{ez} and
\texttt{DiceKriging}, as the former is predominately visualisation of
experimental data and the latter is about the analysis of computer
experiments in addition to belonging to the same suite of packages as
\texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015).

@@

Factorial experiments offers a challenge in allocating the treatment
factors to experimental units where the full set of factorial treatments
cannot be administered (and replicated) and/or the experimental units
have some grouping structure. The effort to address this challenge is
reflected by the number of packages that focus on the construction of
factorial designs as described next.

The \texttt{DoE.base} package (Grömping 2018) can construct full
factorial and (regular and irregular) orthogonal array designs via
\texttt{fac.design()} and \texttt{oa.design()}, respectively. The
\texttt{FrF2} package (Grömping 2014) constructs regular fractional
2-level factorial designs using \texttt{FrF2()} and \texttt{FrF2Large()}
with the latter for large designs. The \texttt{conf.design} package
(Venables 2013) constructs symmetric confounded factorial designs via
\texttt{conf.design()}, and \texttt{BHH2} package (Barrios 2016)
generates full or fractional 2-level factorial design matrix via
\texttt{ffDesMatrix()}.

While the argument names and underlying algorithms of these functions to
generate factorial designs differ, it generally requires the users to
input the number of:

\begin{itemize}
\tightlist
\item
  treatment factors,
\item
  experimental runs (or replications), and
\item
  levels for each factor if the design is allowed to vary in the number
  of levels.
\end{itemize}

The output of the design is either a special class of list (e.g.~the
\texttt{design} class for \texttt{DoE.base} and \texttt{FrF2} or a
\texttt{data.frame} for \texttt{conf.design}) or a matrix for
\texttt{BHH2} such that an element or column corresponds to a treatment
factor with each value corresponding to one experimental run. The
treatment factor are generally assigned pseudo names (e.g.~letters of
the alphabet) or an argument exists for users to input treatment names
as a character vector.

Some forms of factorial designs are known by other names,
e.g.~\emph{response surface designs} are factorial designs where the
treatment factors are discrete levels of continuous variables. Two types
of response surface designs can be constructed using the \texttt{rsm}
package (Lenth 2009): Box-Behnken design (Box and Behnken 1960) and
central-composite designs (Box and Wilson 1951) via functions
\texttt{bbd()} and \texttt{ccd()}, respectively, where the minimum
required input is the number of factors. Another form of factorial
designs is \emph{saturated designs} where typically higher-order
interaction effects of treatment factors are confounded with the main
effects. Plankett-Burnman designs (Plackett and Burman 1946) is one type
of saturated design which can be generated by the function \texttt{pb()}
in the \texttt{FrF2} package, where the user provides the number of
experimental runs and the number of treatment factors.

\hypertarget{the-case-of-recipe-designs}{%
\subsection{The case of recipe
designs}\label{the-case-of-recipe-designs}}

The \texttt{agricolae} package (de Mendiburu 2021) is the prime example
of constructing designs based on a set of so-called ``recipe
functions'', where each function corresponds to a single class of
experimental design. E.g. \texttt{design.crd()}, \texttt{design.rcbd()}
and \texttt{design.split()} construct a completely randomised design,
randomised complete block design and split-plot design, respectively.
Users typically supply the treatment labels (or number of treatments in
the case of \texttt{design.split()}) and the number of replications as
argument to these functions and the output is a list with one element
corresponding to a \texttt{data.frame} that contains the design in a
table such that the row corresponds to the experimental run and the
columns corresponds to the experimental variables (we refer to this
format simply as ``table format'' henceforth).

The use of recipe functions is not limited to classical experimental
designs; the \texttt{AlgDesign} (Wheeler 2022) package offers three
primary functions for generating optimal designs: \texttt{optBlock()},
\texttt{optFederov()} and \texttt{optMonteCarlo()}. These functions in
general require a data and formula in terms of the supplied data with
the choice of criterion (e.g.~D-criterion) with the output as a list
with one element corresponding to the design in a table format. The
difference between these functions lies in the underlying search
strategy for the optimal design and the name of the function is a
surrogate for the search algorithm.

Computer experiments, which generally involve space filling designs, are
implemented in packages like \texttt{lhs} (Carnell 2022) and
\texttt{DiceDesign} (Dupuy, Helbert, and Franco 2015). For the
\texttt{lhs} package, functions like \texttt{randomLHS()},
\texttt{optimumLHS()}, and \texttt{maximinLHS()}, require users to
specify the sample size (\(n\)) and the number of variables (\(p\)) then
it generates a Latin hypercube sample (McKay, Beckman, and Conover
1979), based on different optimisation schemes (in this case, random, S
optimal and maxmin criteria, respectively; see package documentation for
more details). Similarly for the \texttt{DiceDesign}, there is a
comprehensive list of space-filling designs such as
\texttt{dmaxDesign()}, \texttt{lhsDesign()}, and \texttt{wspDesign()}
with input of \(n\) and \(p\) as before (among extra parameters for
some) that implement algorithms that either maximise the entropy (Shewry
and Wynn 1987), produce a random Latin hypercube sample, and use the WSP
algorithm (Santiago, Claeys-Bruno, and Sergent 2012), respectively.
These designs all output a \(n \times p\) matrix with values between 0
and 1. Again, these functions employ a recipe style where each functions
have a name that correspond to a certain search strategy in generating
the experimental design.

\hypertarget{the-case-of-augmenting-designs}{%
\subsection{The case of augmenting
designs}\label{the-case-of-augmenting-designs}}

Some functions in DoE packages require input of existing experimental
designs to produce new designs. For example, the \texttt{DoE.base}
package contain some experimental functions, \texttt{cross.design()} and
\texttt{param.design()}, to combine designs with the former taking a
Cartesian product of the input designs while the latter using a Taguchi
style (Taguchi 1986) to aggregate the designs with the inner and outer
arrays. The \texttt{lhs} package contain the function that
\texttt{augmentLHS()} to add additional samples to existing Latin
hypercube sample.

Another class of augmenting design is \emph{sequential design} (also
called \emph{adaptive sampling}), which is best represented by
\texttt{tgp} (Gramacy and Taddy 2010) -- these require prior
information, which are used to inform the next experimental design using
\texttt{tgp.design()} and \texttt{dopt.gp()}. The user is required to
supply candidate samples to sub-sample from and a model or a prior
experimental design. Follow-up experiments, which can also be classified
as a sequential design, is implemented by \texttt{BsMD} (Barrios 2020)
using a model discriminant approach using \texttt{MD()}.

@@

Through the exploratory data analysis of three data sources (package
download logs, package metadata and CRAN task views) outlined in Section
\ref{data}, we have observed in Section \ref{eda} that the the total
download of DoE packages are concentrated only on a handful of
R-packages although these represent a diverse set in comparison with
other CRAN task views. Furthermore, the data suggests that the field of
experimental design is the least collaborative field.

There are a number of limitations and short coming in our exploratory
data analysis. First, CRAN task views are volunteer maintained so some
experimental design packages may not be included in the DoE packages.
Second, we only use the RStudio CRAN mirror download, which may bias our
observations. Third, our analysis are limited to R-packages alone and
many practioners may indeed be using other methods to construct
experimental designs. Finally, all our statements should be treated as
speculative rather than conclusive; the data are all observational so no
conclusive, generalisable statement are possible. Regardless, the data
driven nature of our analysis give an objective insight into the field
of experimental design.

The interface design (discussed in Section \ref{design}) reveals that
the most widely used DoE packages generally have functions that 1) focus
on certain aspects of experimental design (e.g.~factorial structure or
augmenting design), 2) is a recipe format (i.e.~the name of the function
is a surrogate to a single class of the design or optimal search
algorithm) and 3) context is often a second-thought -- many inputs a
single integer corresponding to the number of factors, levels, or
experimental runs (or sample size). The function will often assign
pseudo factor names or there is an optional argument to input a
character vector that corresponds to the factor names. These interface
designs require users to have processed the experiment in statistical
terms (often stripping the experimental context away) and
simultaneously, users must make a choice of the generation mechanism by
selecting an appropriate function. Arguably, the current dominant
interface designs are not aligned with the way practitioners cognitively
design experiments. Often the critical part of designing an experiment
is understanding the experimental structure and the experimental context
can govern or guide the choice of the algorithm to allocate treatments.
In addition, the nature of recipe functions can obscure the
understanding and relation of the designs (e.g.~how do you go from an
unstructured factorial design to split plot design?). Each new method to
generating an experimental design appear to correspond to a totally new
function and any intermediate results are often not easily accessible.
These may be contributing factors to why developers often work in silos
in the field of experimental design.

A consistent cognitive interface design that leverage existing
developments and can be easily extensible for new methods will
conceivably be of great help to practitioners. Some efforts to this end
is seen in \texttt{DoE.wrapper} (Grömping 2020), which contain wrapper
recipe functions to other DoE packages such as \texttt{lhs},
\texttt{AlgDesign} and \texttt{FrF2} (see Figure
\ref{fig:plot-doe-network}) and is also the subject of the developmental
package \texttt{edibble} (Tanaka 2021). Undoubtedly, no single developer
or package can cater for all experimental design needs so any unifying
interface should consider how other developers can contribute or add
their methods. Future research could benefit from further exploratory
data analysis, expanding the study to beyond R-packages, and discussing
other aspects of interface designs.

@@

\leavevmode\vadjust pre{\hypertarget{ref-DoE.wrapper}{}}%
---------. 2020. \emph{DoE.wrapper: Wrapper Package for Design of
Experiments Functionality}.
\url{https://CRAN.R-project.org/package=DoE.wrapper}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-McKay1979-aw}{}}%
McKay, M D, R J Beckman, and W J Conover. 1979. {``Comparison of Three
Methods for Selecting Values of Input Variables in the Analysis of
Output from a Computer Code.''} \emph{Technometrics: A Journal of
Statistics for the Physical, Chemical, and Engineering Sciences} 21 (2):
239--45. \url{https://doi.org/10.1080/00401706.1979.10489755}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Santiago2012-ui}{}}%
Santiago, J, M Claeys-Bruno, and M Sergent. 2012. {``Construction of
Space-Filling Designs Using {WSP} Algorithm for High Dimensional
Spaces.''} \emph{Chemometrics and Intelligent Laboratory Systems} 113
(April): 26--31. \url{https://doi.org/10.1016/j.chemolab.2011.06.003}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Shewry1987-jn}{}}%
Shewry, M C, and H P Wynn. 1987. {``Maximum Entropy Sampling.''}
\emph{Journal of Applied Statistics} 14 (2): 165--70.
\url{https://doi.org/10.1080/02664768700000020}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Taguchi1986-ut}{}}%
Taguchi, Genichi. 1986. \emph{Introduction to Quality Engineering:
Designing Quality into Products and Processes}. Quality Resources.

\leavevmode\vadjust pre{\hypertarget{ref-edibble}{}}%
Tanaka, Emi. 2021. {``{edibble R-package}.''} \emph{GitHub Repository}.
\url{https://github.com/emitanaka/edibble}; GitHub.
